Geometry of Ranked Nearest Neighbour Interchange Space of Phylogenetic Trees

### 2019-12-11 rnni_convexity.tex
RNNI diameter proof is missing!

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{float}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{todonotes}
\usepackage{enumitem}
\usepackage[misc]{ifsym}
\usepackage[foot]{amsaddr}
\usepackage[hidelinks]{hyperref}
\usepackage{csquotes}
\usepackage[style=authoryear,ibidtracker=false,uniquename=false,giveninits=true,terseinits=true,maxbibnames=5,backend=biber]{biblatex}
\renewbibmacro{in:}{}
\addbibresource{rnni_convexity.bib}

@@

In this section we establish diameter and radius of the $\rnni$ graph.
In Theorem 7 of \autocite{Gavryushkin2018-ol} an upper bound for the diameter is given, but we are able to give the exact diameter using the results of Section~\ref{section:caterpillar_convex}.
In the following we prove the exact diameter of $\rnni$ before we investigate the radius of $\rnni$.

\begin{lemma}
There are two caterpillar trees in $\rnni$ with distance $\frac{(n-1)(n-2)}{2}$.
\label{lemma:caterpillar_diameter}
\end{lemma}

\begin{proof}
Let $T = [1,2,3,\ldots,n]$ and $R = [n-1,n,n-2,n-3, \ldots, 1]$ be two caterpillar trees.
According to Theorem~\ref{thm:caterpillar_convex}, there is a caterpillar path of length $d(T,R)$ between $T$ and $R$.
We can compute a shortest caterpillar path by using $\csort$ as explained in Section~\ref{section:caterpillar_convex}.
When running this algorithm to compute a path from $T$ to $R$, each taxon $i$ ($i = n, \ldots, 3$) swaps $i-2$ times with its upper neighbour.
Therefore, the length of a shortest caterpillar path, and therefore the distance between $T$ and $R$, is $\sum\limits_{i=3}^{n}(i-2)  = \frac{(n-1)(n-2)}{2}$.
\end{proof}

@@

>>>>

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{float}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{todonotes}
\usepackage{enumitem}
\usepackage[misc]{ifsym}
\usepackage[foot]{amsaddr}
\usepackage[hidelinks]{hyperref}
\usepackage{csquotes}
\usepackage[style=authoryear,ibidtracker=false,uniquename=false,giveninits=true,terseinits=true,maxbibnames=5,backend=biber]{biblatex}
\renewbibmacro{in:}{}
\addbibresource{rnni_convexity.bib}

@@

\textcite[Theorem~7]{Gavryushkin2018-ol} gave an upper bound for the diameter of $\rnni$ space.
In this section we improve that result and calculate the exact diameter and radius of the $\rnni$ space.
To achieve this we will employ our results on caterpillar trees from the previous sections.

The results of previous section imply that $d([1,2,3,\ldots,n], [n-1,n,n-2,n-3, \ldots, 1]) = \frac{(n-1)(n-2)}{2}$, which is the maximum possible distance between caterpillar trees.

@@

\proof
\todo{Why exactly is this true for arbitrary, not necessarily caterpillar, trees?}
\endproof
